---
title: "Feather Performance Analysis"
author: "Edgar Ruiz"
date: "June, 3 2016"
output: html_document
---

### Introduction

The set is made up of 32 CSV files. Each file contain 80 variables. The smallest file is 50 megabytes and the largest is a bit under 2 gigabytes. Each file is 50 megabytes larger than the previous one, or around 86,000 rows. Details of the files sizes are available on Appendix A

The steps for each experiment are:

1. Read CSV file using 'read.csv'
2. Read CSV file using 'fread'
3. Write CSV file using 'write.csv'
4. Write Feather file
5. Read the Feather file

Each experiment takes longer than the previous one, so the entire set of 32 files takes 2 hours to run on my laptop, which is how long this 'R Markdown' report took to complete.

The code will also save the results into two feather files which are used to write up the analysis.

#### Tidy Dataset

```{r, eval=FALSE}
library(feather)
library(data.table)
library(stringr)
library(ggplot2)

csv_files <- list.files(path="csvfiles/")
entry_dataset <- NULL
for(j in 1:length(csv_files)){
  current_file <- csv_files[j]
  csv_file  <- paste("csvfiles/",current_file, sep="")
  csv_new <- "new.csv"
  feather_file <- str_replace_all(csv_file, "csv","feather")
  #-------------------- Reading CSV file using read.csv & recording results to data frame ----------------
  entry_dataset <- rbind(entry_dataset,data.frame(testfile=csv_file, tech="csv",step="read",elapsed=system.time(original_csv <- read.csv(csv_file))[3] ,rows=nrow(original_csv), columns=ncol(original_csv), filesize=file.info(csv_file)$size, memsize=object.size(original_csv)[1]))
  #-------------------- Write CSV file & recording results to data frame ---------------------------------
  entry_dataset <- rbind(entry_dataset,data.frame(testfile=csv_file, tech="csv",step="write",elapsed=system.time(write.csv(original_csv,csv_new))[3] ,rows=nrow(original_csv), columns=ncol(original_csv), filesize=file.info(csv_new)$size, memsize=0))
  #-------------------- Reading CSV file using 'fread' & recording results to data frame -----------------
  entry_dataset <- rbind(entry_dataset,data.frame(testfile=csv_file, tech="feather",step="write",elapsed=system.time(write_feather(original_csv, feather_file))[3] ,rows=nrow(original_csv), columns=ncol(original_csv), filesize=file.info(feather_file)$size, memsize=0))
  rm(original_csv)
   #-------------- Writting data frame using Feather& recording results to data frame --------------------
  entry_dataset <- rbind(entry_dataset,data.frame(testfile=csv_file, tech="feather",step="read",elapsed=system.time(new_feather<- read_feather(feather_file))[3] ,rows=nrow(new_feather), columns=ncol(new_feather), filesize=file.info(feather_file)$size, memsize=object.size(new_feather)[1]))
  rm(new_feather)
  #-------------------- Reading Feather file & recording results to data frame -----------------
  entry_dataset <- rbind(entry_dataset,data.frame(testfile=csv_file, tech="dt",step="read",elapsed=system.time(dt_csv <- fread(csv_file))[3] ,rows=nrow(dt_csv), columns=ncol(dt_csv), filesize=0, memsize=object.size(dt_csv)[1]))
  rm(dt_csv)
}
csv_files <- entry_dataset[entry_dataset$tech=="csv" & entry_dataset$step=="read",c("rows","filesize")]
colnames(csv_files) <- c("rows","original_filesize")
#----------------- Adding the original CSV file size to ease comparison -------------------------
entry_dataset <- merge(x=entry_dataset, y=csv_files,  by="rows")
write_feather(entry_dataset, "results.feather")
print(structure(entry_dataset))
```

```{r, include=FALSE}
#-----------  The same code repeats to supress fread messages  -----------------------
library(feather)
library(data.table)
library(stringr)
library(ggplot2)

csv_files <- list.files(path="csvfiles/")
entry_dataset <- NULL
for(j in 1:length(csv_files)){
  current_file <- csv_files[j]
  csv_file  <- paste("csvfiles/",current_file, sep="")
  csv_new <- "new.csv"
  feather_file <- str_replace_all(csv_file, "csv","feather")
  #-------------------- Reading CSV file using read.csv & recording results to data frame ----------------
  entry_dataset <- rbind(entry_dataset,data.frame(testfile=csv_file, tech="csv",step="read",elapsed=system.time(original_csv <- read.csv(csv_file))[3] ,rows=nrow(original_csv), columns=ncol(original_csv), filesize=file.info(csv_file)$size, memsize=object.size(original_csv)[1]))
  #-------------------- Write CSV file & recording results to data frame ---------------------------------
  entry_dataset <- rbind(entry_dataset,data.frame(testfile=csv_file, tech="csv",step="write",elapsed=system.time(write.csv(original_csv,csv_new))[3] ,rows=nrow(original_csv), columns=ncol(original_csv), filesize=file.info(csv_new)$size, memsize=0))
  #-------------------- Reading CSV file using 'fread' & recording results to data frame -----------------
  entry_dataset <- rbind(entry_dataset,data.frame(testfile=csv_file, tech="feather",step="write",elapsed=system.time(write_feather(original_csv, feather_file))[3] ,rows=nrow(original_csv), columns=ncol(original_csv), filesize=file.info(feather_file)$size, memsize=0))
  rm(original_csv)
   #-------------- Writting data frame using Feather& recording results to data frame --------------------
  entry_dataset <- rbind(entry_dataset,data.frame(testfile=csv_file, tech="feather",step="read",elapsed=system.time(new_feather<- read_feather(feather_file))[3] ,rows=nrow(new_feather), columns=ncol(new_feather), filesize=file.info(feather_file)$size, memsize=object.size(new_feather)[1]))
  rm(new_feather)
  #-------------------- Reading Feather file & recording results to data frame -----------------
  entry_dataset <- rbind(entry_dataset,data.frame(testfile=csv_file, tech="dt",step="read",elapsed=system.time(dt_csv <- fread(csv_file))[3] ,rows=nrow(dt_csv), columns=ncol(dt_csv), filesize=0, memsize=object.size(dt_csv)[1]))
  rm(dt_csv)
}
csv_files <- entry_dataset[entry_dataset$tech=="csv" & entry_dataset$step=="read",c("rows","filesize")]
colnames(csv_files) <- c("rows","original_filesize")
#----------------- Adding the original CSV file size to ease comparison -------------------------
entry_dataset <- merge(x=entry_dataset, y=csv_files,  by="rows")
write_feather(entry_dataset, "results.feather")
print(head(entry_dataset))
```


### Comparison Dataset

```{r}
csv <- entry_dataset[entry_dataset$tech=="csv" & entry_dataset$step=="read",]
colnames(csv) <- paste("csv_", names(csv), sep="")
dt <- entry_dataset[entry_dataset$tech=="dt" & entry_dataset$step=="read",]
colnames(dt) <- paste("dt_", names(dt), sep="")
feather <- entry_dataset[entry_dataset$tech=="feather" & entry_dataset$step=="read",]
colnames(feather) <- paste("feather_", names(feather), sep="")
step_read <- merge(x=csv, y=dt, by.x="csv_rows", by.y="dt_rows")
step_read <- merge(x=step_read, y=feather, by.x="csv_rows", by.y = "feather_rows")
step_read$cf_elapsed <- step_read$csv_elapsed / step_read$feather_elapsed
step_read$df_elapsed <- step_read$dt_elapsed / step_read$feather_elapsed
write_feather(step_read, "comparison.feather")
print(head(step_read))
```

## Appendix

### Appendix A
```{r}
csv_files <- file.info(paste("csvfiles/",list.files(path="csvfiles/"),sep=""))
print(data.frame(FileName = rownames(csv_files), FileSize_Bytes = prettyNum(csv_files$size, big.mark = ",")))
```

